### 2008-12-19 talk/ecoop2009/intro.tex
remove the jitstrategy section and integrate its contents into the introduction.

This section will give a high-level overview of how the JIT-generation process
works. More details will be given in subsequent sections.

- write an interpreter
- add some hints to the interpreter

translation time:
- start the translation process, flow graphs
- run a binding-time analysis using the hints, to figure out ...
- turn the flow graphs into rainbow bytecode
- turn the flow graphs of the rainbow interpreter and the original flow graphs
into .NET code

runtime:
when the interpreter is started and a function should be interpreted:
 - the generated JIT starts compiling the function 
 - it produces .NET bytecode until it does not have enough information to
 continue. then it stops the compilation and puts a callback into the compiler
 into the generated bytecode.
 - at this point the .NET bytecode that is produced so far is executed.
 - when the callback into the compiler is hit, execution stops and compilation
 starts again
 - afterwards execution continues after the callback (running the code that was
 newly produced)

>>>>

One of the most important aspects that PyPy's translation toolchain can weave
in is the \emph{automatic generation of a JIT compiler}.  This section will
give a high-level overview of how the JIT-generation process works. More
details will be given in subsequent sections.

The first step is to write an interpreter for the chosen language.  Since it
must be fed to the translation toolchain, the interpreter has to be written in
RPython.  Then, to get best performances, we need to add few manual
annotations to the interpreter, in order to teach the JIT generator which
informations are important to know at compile-time.  Annotations are inserted
as \emph{hints}, as described in section \ref{sec:hints}.

It is important to distinguish between three distinct phases of execution:

\begin{enumerate}
\item \textbf{Translation-time}: when the translation toolchain runs and the
  JIT compiler is automatically generated from the interpreter.  The result is
  an executable that can be used to run programs written in the chosen
  language.
\item \textbf{Compile-time}: when the JIT compiler runs, generating executable
  code on the fly.
\item \textbf{Runtime}: when the code generated at compile-time is executed.
\end{enumerate}

Note that in this schema translation-time happens only once, on the
developer's machine.  By contrast, compile-time and runtime happens every time
the user wants to run some program.

\subsection{Effective JIT compilation of dynamic languages}

Generating efficient compilers for dynamic languages is hard.  Since these
languages are dynamically typed, usually the compiler does not have enough
information to produce efficient code, but instead it has to insert a lot of
runtime checks to select the appropriate implementation for each operation.

By emitting code at runtime, JIT compilers can exploit some extra knowledge
compared to traditional static compilers.  However, we need to take special
care to choose a strategy for JIT compilation that lets the compiler to take
the best of this advantage.

Most JIT compilers for dynamic languages around (such as IronPython, Jython,
JRuby \anto{XXX insert some reference}) compile code at the method
granularity.  If on one hand they can exploit some of the knowledge gathered
at runtime (e.g. the types of method parameters), on the other hand they can
do little to optimize most of the operations inside, because their behaviour
depends on informations that are not available at compile-time, because
e.g. the global state of the program can change at runtime. \anto{e.g., we can
  add/remove methods to classes, etc. Should we insert some example here?}

JIT compilers generated by PyPy solve this problem by delaying the compilation
until they know all the informations needed to generate efficient code.  If at
some point the JIT compiler does not know about something it needs, it
generates a callback into itself.

Later, when the generated code is executed, the callback is hit and the JIT
compiler is restarted again.  At this point, the JIT knows exactly the state
of the program and can exploit all this extra knowledge to generate highly
efficient code.  Finally, the old code is patched and linked to the newly
generated code, so that the next time the JIT compiler will not be invoked
again.  As a result, \textbf{runtime and compile-time are continuously
  intermixed}.

Modifying the old code to link the newly generated one is very challenging on
.NET, as the framework does not offer any primitive to do this.  Section
\ref{sec:clibackend} explains how it is possible to simulate this behaviour.
